\hspace{1cm} Do syntactic and semantic similarity lead to interference effects? \newline Evidence from self-paced reading and event-related potentials using German

### 2024-07-15 paper/main.tex
Updated via the Open Science Framework

Self-paced reading times were analyzed with Bayesian linear mixed models in R \citep{r} with log-normal likelihood, using the brms package \citep{brms}. The critical word for analyses was the verb \textit{trank} `drank' in (\ref{ex:materials}), constituting the retrieval site. Because previous work found effects in the pre-critical and post-critical region, reading times of the pre-critical word (\textit{regelmäßig} `regularly' in \ref{ex:materials}) and the spill-over region (\textit{um} `in.order.to' in \ref{ex:materials}) were analyzed. 
The models included fixed effects for syntactic interference (high +0.5, low -0.5), semantic interference (high +0.5, low -0.5) and their interaction.  Varying intercepts for participants and items were included; varying slopes were excluded for reasons explained below.

\paragraph{Bayes factors for model comparison}

In order to quantify the uncertainty on the parameters of interest, we report 95\% credible intervals.  
These intervals represent the range over which we can be 95\% certain that the values of the parameter lies, given the statistical model and the data. However, formal hypothesis testing cannot be carried out without a likelihood ratio test that compares two alternative models \parencite{schad_etal_2022_BF,Royall}. For this reason, formal hypothesis tests for the presence or absence of the effects of interest were carried out using Bayes factors. Because Bayes factors can be very sensitive to prior specifications on the target parameter being tested \parencite{schad_etal_2022_BF}, 
we report a sensitivity analysis using a range of prior specifications for the relevant parameters (see Table \ref{tab:spr_priors}). The priors assume a priori effect sizes for the main effects of syntactic and semantic interference, and their interaction, ranging from 0-8 ms, 0-40 ms, or 0-81 ms. These priors are based on the observed sign and sizes of effects and uncertainties in previous reading studies that use the present design \parencite{vandyke07,mertzen}, and on meta-analyses relating to interference effects \parencite{jaeger_etal_2017}. For the syntactic and semantic interference effects, we use directional priors so that any evidence for the effect of interest from Bayes factors can be interpreted as evidence of replication success \parencite{verhagen2014bayesian}.

@@

\begin{table}[h]
    \caption{Priors used for the analysis of the self-paced reading time data. Priors on the fixed-effects slope parameters were truncated to a lower bound of zero, i.e., only positive values were assumed to be possible a priori. This corresponds to the existing evidence in the literature that high interference conditions lead to longer reading times than low interference conditions. The standard deviation of the fixed-effects slope priors was varied in order to conduct a Bayes factor sensitivity analysis following \citet{schad_etal_2022_BF}. See the corresponding assumed a priori range of the difference between reading times under high vs.\ low interference on the millisecond scale.}
    \label{tab:spr_priors}
    \centering
     \begin{tabular}{lllr}
    \toprule
    Parameter&Prior & &Assumed Range in ms\\
    \midrule
  Intercept & Normal(6, 0.6)& & [125, 1308]\\
  \cmidrule{2-4}
  \multirow{3}{1cm}{slope} & Normal(0, 0.01) & \hspace{-1em}\rdelim\}{3}{*}[lower bound: 0] & [0, 8]\\
  &  Normal(0, 0.05)& & [0, 40]\\
  & Normal(0, 0.1) & &[ 0, 81]\\
  \cmidrule{2-4}
  sigma & Normal(0, 0.5)&  \hspace{-1em}\rdelim\}{2}{*}[lower bound: 0]&\\
  SD & Normal(0, 0.1)& &\\
  \bottomrule
  \end{tabular}
  \end{table}

@@

Turning now to the estimates of the main effects and interaction parameters in the pooled data (Experiments 1a and 1b combined), Figure \ref{fig:spr_posteriors} shows that the pre-critical, critical and spill-over regions showed very similar posterior distributions for the parameters. For syntactic interference, they were close to zero in all regions. In contrast, for semantic interference, the estimates ranged between [8,14] ms (pre-critical), [4, 12] ms (critical), and [5, 11] ms (spill-over region). The posteriors for the interaction were consistent with an interaction effect. The interaction was the  largest in the critical region ([7, 22 ms]).

@@

Was there any evidence for the effects estimated above? The Bayes factor analysis is presented in Figure \ref{fig:spr_bfs}. The analyses use all the data from Experiments 1a and 1b. We found extremely strong evidence for the semantic interference effect in the pre-critical, critical and spill-over regions (BF$_{10}$ > 200). No evidence was found for syntactic interference in any of the investigated regions; in fact, there is strong evidence against the syntactic interference effect in all regions and under all the priors used (BF$_{10}$ < 1). Evidence for the interaction was only found in the critical region. Here, the evidence for the interaction was strong (BF$_{10}$ = 40) when the prior on the parameter was N+(0, 0.01), i.e., a small a priori effect size ranging from  [0, 8] ms; the evidence for the interaction was much weaker (BF$_{10}$ = 1.8) when the prior was N+(0,0.1), i.e., a larger a priori effect size ranging from [0, 81] ms. In other words, only if we assume a priori that the effect size for the interaction is relatively small, then there is strong evidence for this effect.

@@

\subsection{Discussion}
We presented the to-date largest-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation. Reading times at the critical verb showed clear evidence for semantic interference, but evidence against syntactic interference. The semantic interference effect was already present at the pre-critical region; i.e., semantic interference occurred even before the retrieval event could begin. The effects in the pre-critical region are discussed next.

@@

@@

\subsubsection{Comparison with previous findings}
The results of the present study stand in contrast to previous findings \citep{vandyke07, mertzen}. The present study is the only one to find semantic and no syntactic interference effects. The difference in observed effects could be explained by differences in statistical power or by methodological differences. The present study has higher statistical power than the previous studies. \citet{vandyke07} had 35 - 40 participants and 36 - 48 items in each of her three experiments, respectively. \textcite{mertzen} tested 61 English speakers and 121 German speakers with 40 items each. In contrast, we tested 774 participants with at least 60 items each (a subset of 160 participants read all 120 items). Therefore, the present study is the study with the highest power to date using this 2 $\times$ 2 interference design. Consequently, the estimates from the present study are the most precise ones so far for this design. The smaller effect size in the present study is a consequence of higher power, and the magnitude of the effects observed is comparable to that of other large-scale reading studies \citep{nicenboim} and meta-analyses on interference \citep{jaeger_etal_2017}.

So, why did we find a semantic interference effect in all regions, whereas \citet{mertzen} only found one in the pre-critical region? Apart from \textcolor{blue}{the additional distractor and} the methodological difference (eye-tracking vs.\ self-paced reading), a plausible  explanation for the divergence is lower power in \citeauthor{mertzen}'s (\citeyear{mertzen}) study. While they already had considerably more participants and items than many psycholinguistic studies, in the present study the number of participants was much higher than in theirs (774 vs.\ 121) and the number of items was also higher (60/120 vs.\ 40). 
We will revisit the absence of syntactic interference in our data in the general discussion.

@@

\begin{table}[!htbp]
    \caption{Relatively informative priors for the analysis of the event-related potentials \citep{nicenboim_stats}. Priors on the slope parameters were truncated to an upper bound of zero, i.e., only negative values. This corresponds to more negative brain responses to high vs.\ low interference conditions in line with the literature \citep{lee_garnsey, martinetal2014, schoknecht2022, vasishth_drenhaus_2011}. The directional priors also serve to allow us to quantify the extent to which a predicted effect replicates; the Bayes factors can be interpreted as indicating the evidence supporting or not supporting replication success \citep{verhagen2014bayesian}.
    The standard deviations of the slope priors were varied in order to conduct a Bayes factor sensitivity analysis. See the corresponding assumed a priori range of the difference between high vs.\ low interference.}
    \label{tab:eeg_priors}
    \centering
    \begin{tabular}{lllr}
    \toprule
    Parameter&Prior & &Assumed Range ($\mu V$)\\
    \midrule
  Intercept & Normal(2, 5)& & [-8, 12]\\
  \cmidrule{2-4}
  \multirow{4}{1cm}{slope} & Normal(0, 0.1) &\hspace{-1em}\rdelim\}{4}{*}[upper bound: 0]& [-0.2, 0]\\
  & Normal(0, 0.5)& & [-1, 0]\\
  & Normal(0, 1) && [-2, 0]\\
  & Normal(0, 5) && [-11, 0]\\
  \cmidrule{2-4}
  sigma & Normal(10, 5)& \hspace{-1em}\rdelim\}{2}{*}[lower bound: 0]\\
  SD & Normal(0, 2)&\\
 \textcolor{blue}{correlations} & \textcolor{blue}{LKJcorr(2)} & & \\
    \bottomrule
    \end{tabular}
\end{table}

Bayesian models were run with four chains and 20,000 iterations of which 2,000 were used as warm-up phase \citep{schad_etal_2022_BF}. For the calculation of Bayes factors and the corresponding sensitivity analysis, we defined a range of priors on the parameters of interest, assuming a range of effect sizes \citep{nicenboim_stats,schad_etal_2022_BF}. We chose slope priors which assume a wide range of effect sizes from $-0.2$ to $-11$\,$\mu V$. However, the more realistic a priori effect sizes probably range from $-0.2$ to at most $-2$\,$\mu V$. This assessment is based on our expectation that the interference effect investigated here, being a subtle manipulation, is probably smaller in magnitude than the well-established  effect of cloze probability on the N400 amplitude analyzed by \textcite{nicenboim_stats}, which ranges from $0$ to $5$\,$\mu V$ \parencite[see section 5.2.5][]{nicenboim_stats}.

@@

Figure \ref{fig:erp_all} shows the 
grand average ERPs elicited by the critical verb for all conditions. Table \ref{tab:eeg_mod} shows the estimates for the brain activity in the spatio-temporal N400 window from the linear mixed model with the truncated prior Normal$_-$(0, 0.5) which assumed more negative ERPs for high vs.\ low interference, with a difference possibly as large as -1\,$\mu V$. The 95\% credible intervals for the fixed effects are consistent with effects of syntactic interference, semantic interference and their interaction. The estimate of the semantic interference effect was the largest and its 95\% credible interval was the furthest from zero (\textcolor{blue}{mean: -0.32\,$\mu V$, 95\% CrI = [-0.59, -0.10]\,$\mu V$}, for the posteriors, see Figure \ref{fig:eeg_posteriors}). The estimates of syntactic interference and the interaction were both about half the size of the semantic effect (see Table \ref{tab:eeg_mod} and Figure \ref{fig:eeg_posteriors}).

@@

\begin{table}[H]
    \caption{Results from the Bayesian linear mixed model analyzing the averaged brain response from 12 centro-parietal electrodes in the time window 300 to 500 ms post critical verb onset. The predictor baseline represents the averaged activity from the 200 ms pre-verb onset and was included in the model as a covariate instead of traditional baseline correction \citep[see][]{alday2019}.}
    \label{tab:eeg_mod}
    \centering
    \begin{tabular}{lrrr}
    \toprule
    & Estimate & Est.Error & 95\% CrI  \\
    \midrule
Intercept &  -0.70 & 0.19 &  [-1.06, -0.34] \\
baseline  &  -0.22 & 0.01 &  [-0.24, -0.20] \\
syntactic &  \textcolor{blue}{-0.16}& \textcolor{blue}{0.08} &  \textcolor{blue}{[-0.33, -0.02]} \\
semantic  &  \textcolor{blue}{-0.30} & \textcolor{blue}{0.10} &  \textcolor{blue}{[-0.50, -0.11]} \\
interaction& \textcolor{blue}{-0.14} & \textcolor{blue}{0.11} &  \textcolor{blue}{[-0.40, -0.01]} \\
    \bottomrule
    \end{tabular}
\end{table}

\begin{figure}[H]
    \caption{Posteriors of the syntactic and semantic interference effects and their interaction for the event-related potentials elicited by the critical verb in the spatio-temporal N400 window. Blue vertical lines represent the median and shaded areas are 80\,\% intervals.}
    \label{fig:eeg_posteriors}
    \centering
    \includegraphics[width=0.7\textwidth]{images/posteriors_eeg.jpg}
\end{figure}

@@

@@

In the remainder of this section, we discuss two important issues that relate to previous experiments that investigated syntactic and semantic interference. The first is the role of encoding vs.\ retrieval interference, and the second is broader implications of our findings in the context of previous related work.

>>>>

Self-paced reading times were analyzed with Bayesian linear mixed models in R \citep{r} with log-normal likelihood, using the brms package \citep{brms}. The critical word for analyses was the verb \textit{trank} `drank' in (\ref{ex:materials}), constituting the retrieval site. Because previous work found effects in the pre-critical and post-critical region, reading times of the pre-critical word (\textit{regelmäßig} `regularly' in \ref{ex:materials}) and the spill-over region (\textit{um} `in.order.to' in \ref{ex:materials}) were analyzed. 
The models included fixed effects for syntactic interference (high +0.5, low -0.5), semantic interference (high +0.5, low -0.5) and their interaction. \textcolor{blue}{In addition to the predictors of interest, we included trial id to account for potential effects of fatigue or adaptation to the task. The trial id of the 100 trials per session was recoded to span from 0 to 1. This was done to bring all fixed effects to a comparable scale which decreases the run time of the models. The models were run with full random effects (varying intercepts and slopes of all fixed effects and their interactions for participants and items).}

\paragraph{Bayes factors for model comparison}\label{BF_analysis_SPR}

In order to quantify the uncertainty on the parameters of interest, we report 95\% credible intervals.  
These intervals represent the range over which we can be 95\% certain that the values of the parameter lies, given the statistical model and the data. However, formal hypothesis testing cannot be carried out without a likelihood ratio test that compares two alternative models \parencite{schad_etal_2022_BF,Royall}. For this reason, formal hypothesis tests for the presence or absence of the effects of interest were carried out using Bayes factors. Because Bayes factors can be very sensitive to prior specifications on the target parameter being tested \parencite{schad_etal_2022_BF}, 
we report a sensitivity analysis using a range of prior specifications for the relevant parameters (see Table \ref{tab:spr_priors}). The priors assume a priori effect sizes for the main effects of syntactic and semantic interference, and their interaction, \textcolor{blue}{ranging from -8 to 8 ms, -40 to 40 ms, or -81 to 81 ms. These priors are based on the observed sizes of effects and uncertainties in previous reading studies that use the present design \parencite{vandyke07,mertzen}, and on meta-analyses relating to interference effects \parencite{jaeger_etal_2017}.}

@@

\begin{table}[h]
    \caption{Priors used for the analysis of the self-paced reading time data. The standard deviation of the fixed-effects slope priors was varied in order to conduct a Bayes factor sensitivity analysis following \citet{schad_etal_2022_BF}. See the corresponding assumed a priori range of the difference between reading times under high vs.\ low interference on the millisecond scale.}
    \label{tab:spr_priors}
    \centering
     \begin{tabular}{llr}
    \toprule
    Parameter&Prior &Assumed Range in ms\\
    \midrule
  Intercept & Normal(6, 0.6)& [125, 1308]\\
  \cmidrule{2-3}
  \multirow{3}{1cm}{slope} & Normal(0, 0.01) & \textcolor{blue}{[-8, 8]}\\
  &  Normal(0, 0.05)&  \textcolor{blue}{[-40, 40]}\\
  & Normal(0, 0.1) & \textcolor{blue}{[-81, 81]}\\
  \cmidrule{2-3}
  sigma & Normal(0, 0.5)&\\
  SD & Normal(0, 0.1) &\\
  \bottomrule
  \end{tabular}
  \end{table}

@@

Turning now to the estimates of the main effects and interaction parameters in the pooled data (Experiments 1a and 1b combined), Figure \ref{fig:spr_posteriors} shows that the pre-critical, critical and spill-over regions showed very similar posterior distributions for the parameters. \textcolor{blue}{The 95 \% credible intervals for syntactic interference included zero in the pre-critical and critical regions. In the spill-over region, it included only positive values ([1, 15] ms). In contrast, for semantic interference, the 95 \% credible intervals were fully positive in all regions; with identical estimates in the pre-critical and spill-over region ([7, 21] ms). The posteriors for the interaction were mostly positive, but the 95 \% credible intervals included zero in all regions. The interaction was the  largest in the critical region ([-3, 31 ms]).}

@@

Was there any evidence for the effects estimated above? The Bayes factor analysis is presented in Figure \ref{fig:spr_bfs}. The analyses use all the data from Experiments 1a and 1b. \textcolor{blue}{We found very to extremely strong evidence for the semantic interference effect in the pre-critical and spill-over region (BF$_{10}$ between 36 and 231). In the critical region, the evidence for semantic interference was inconclusive across the range of investigated priors (BF$_{10}$ between 0.7 and 3.5). There was strong evidence against an effect of syntactic interference under almost all priors in all regions (BF$_{10}$ < 0.5). Only in the spill-over region when small effects between -8 and 8 ms were assumed a priori, there was weak evidence for syntactic interference (BF$_{10}$ = 2.8). Similarly, there was overall evidence against the interaction of syntactic and semantic interference (BF$_{10}$ < 1). Only under the narrowest prior in the critical region, there was very weak evidence for the interaction (BF$_{10}$ = 1.4).}

\textcolor{blue}{All in all, the Bayes factor analysis provided very strong evidence for the semantic interference effect in the pre-critical and spill-over region, but inconclusive results in the critical region. In contrast, the Bayes factor analysis provided either at best weak evidence for syntactic interference and the interaction or even evidence against them.}

@@

\subsection{Discussion}
We presented the to-date largest-sample self-paced reading study that investigated the use of syntactic and semantic features during subject-verb dependency formation. \textcolor{blue}{Reading times at the critical verb were inconclusive for semantic interference and showed evidence against syntactic interference. We will discuss the absence of syntactic interference in our data in the general discussion. The semantic interference effect was already present at the pre-critical region; i.e., semantic interference occurred even before the retrieval event could begin.} The effects in the pre-critical region are discussed next.

@@

\textcolor{blue}{Another possible explanation for the reading time differences starting at the distractor could be a difference in plausibility between the high and low semantic interference conditions. To explore this possibility, we conducted a web-based plausibility judgement experiment on PCIbex  \parencite{pcibex}. Since we were only interested in the potential plausibility difference between the semantic interference conditions, we used a one-factorial design for the plausibility rating experiment investigating semantic interference within the high syntactic interference conditions (see \ref{ex:hisyn}). Forty-four native speakers of German (mean age XX, age range: , w, m), who were recruited over prolific and had not participated in the SPR experiment or norming study, provided ratings on a scale from 1 (absolutely implausible) to 7 (absolutely plausible). The experiment included 40 implausible filler sentences which were created by exchanging nouns between the fillers used in the SPR experiment (e.g., Der Sohn des Kaisers, der den Feldzug gewonnen hatte, plante äußerst strategisch, weil er die Gießkanne an sich reißen wollte., `The son of the emperor who had won the campaign planned extremely strategically because he wanted to take the watering can by force.'). Participants spent approximately 37 minutes to complete the task and received 6.50 \pounds \, as compensation. The results of an ordinal brms model analyzing the ratings are shown in Figure X. Sentences with an inanimate distractor had a slightly higher probability to gain a high rating on the plausibility scale ($\beta$ = XX), i.e., the low semantic interference conditions are more plausible. To investigate whether this plausibility difference is driving the semantic interference effect in our SPR data, we ran an additional model which used the same specifications as described in the Statistical Analyses section with prior N(0,0.5) and additionally, the centered plausibility ratings as a fixed effect. Figure X presents the posteriors from this model. 
}

@@

@@

\begin{table}[!htbp]
    \caption{Relatively informative priors for the analysis of the event-related potentials \citep{nicenboim_stats}. The standard deviations of the slope priors were varied in order to conduct a Bayes factor sensitivity analysis. See the corresponding assumed a priori range of the difference between high vs.\ low interference.}
    \label{tab:eeg_priors}
    \centering
    \begin{tabular}{llr}
    \toprule
    Parameter&Prior &Assumed Range ($\mu V$)\\
    \midrule
  Intercept & Normal(2, 5)& [-8, 12]\\
  \cmidrule{2-3}
  \multirow{4}{1cm}{slope} & Normal(0, 0.1) & [-0.2, 0]\\
  & Normal(0, 0.5)& [-1, 1]\\
  & Normal(0, 1) & [-2, 2]\\
  & Normal(0, 2) & [-4, 4]\\
  \cmidrule{2-3}
  sigma & Normal(10, 5)&\\
  SD & Normal(0, 2)&\\
    \bottomrule
    \end{tabular}
\end{table}

Bayesian models were run with four chains and 20,000 iterations of which 2,000 were used as warm-up phase \citep{schad_etal_2022_BF}. For the calculation of Bayes factors and the corresponding sensitivity analysis, we defined a range of priors on the parameters of interest, assuming a range of effect sizes \citep{nicenboim_stats,schad_etal_2022_BF}. We chose slope priors which assume a wide range of effect sizes from [$-0.2$, 0.2] to [$-4$, 4]\,$\mu V$.

@@

Figure \ref{fig:erp_all} shows the 
grand average ERPs elicited by the critical verb for all conditions. \textcolor{blue}{Figure \ref{fig:eeg_posteriors} shows the estimates for the brain activity in the spatio-temporal windows of the N400 and P600 from the linear mixed models with prior Normal(0, 0.5) which assumed interference effects with a difference possibly as large as -1\,$\mu V$. The 95\% credible intervals of the semantic interference effect showed more negative brain responses for high vs.\ low interference and did not include zero for both the N400 and the P600. The 95\% credible intervals of the syntactic interference effect also showed more negative brain responses for high vs.\ low interference, but the mean estimate was about half the size of the semantic interference estimate, respectively, and the intervals include zero for both the N400 and the P600. The 95\% credible interval of the interaction was centered around zero for the N400. For the P600, it included mostly negative values but crossed zero.}

@@

\begin{figure}[H]
    \caption{Posteriors of the syntactic and semantic interference effects and their interaction for the event-related potentials elicited by the critical verb in the spatio-temporal \textcolor{blue}{windows of the N400 (A) and P600 (B)}. Blue vertical lines represent the median and shaded areas are 80\,\% intervals.}
    \label{fig:eeg_posteriors}
    \centering
    \includegraphics[width=\textwidth]{images/posteriors_eeg.jpg}
\end{figure}

@@

% root mean square deviation

@@

\textcolor{red}{Rewrite this section to reflect the new discussion.} In the remainder of this section, we discuss two important issues that relate to previous experiments that investigated syntactic and semantic interference. The first is the role of encoding vs.\ retrieval interference, and the second is broader implications of our findings in the context of previous related work.

\subsection{Comparison with previous findings}
\textcolor{blue}{The present study used the same design as two previous studies \parencite{vandyke07, mertzen}. Figure \ref{fig:previous_vs_present} A shows the estimated reading times differences from the previous studies and our self-paced reading experiment. Because the \textcite{vandyke07} data was not available to us, we derived estimates for their study from the summary statistics in the published paper. \textcite{mertzen} made their data available. In comparison to the previous estimates, the effects of the present study are relatively small and they are associated with less uncertainty. To see whether there is evidence for the estimated effects, we computed Bayes factors for the previous studies. For \textcite{vandyke07}, we used the Normal-Normal conjugate case to derive posteriors from the summary statistics (mean and SE). Then using these posteriors and a reasonable prior (Normal(0,0.5), a priori assumed effect size in the range of -40 ms to 40 ms), we computed Bayes Factors using the Savage-Dickey method. Bayes factors for \citeauthor{mertzen}'s (\citeyear{mertzen}) data were computed using the Savage-Dickey method from models which we run on their data along to our own analyses described in Statistical Analyses for Experiments 1a and b. The Bayes factors of the effects of interest in self-paced reading times and total fixation times of the present and previous studies under prior Normal(0,0.5) are shown in Figure \ref{fig:previous_vs_present} B.}

\begin{sidewaysfigure}[h]
    \caption{A) Reading time differences in the regions of interest from the previous and present studies using the design by \textcite{vandyke07}. VD07.E1, VD07.E2 and VD07.E3 stand for \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiments 1-3. M23.E and M23.G stand for \citeauthor{mertzen}'s (\citeyear{mertzen}) English and German experiment, respectively. The intervals for their study and the present study are Bayesian 95\% credible intervals. There are less estimates for the pre-critical region because \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1 and 2 did not have a pre-critical region. B) Bayes factors for the effects of interest in for the present and previous studies under prior Normal(0,0.5).}
    \label{fig:previous_vs_present}
    \centering
    \includegraphics[width=0.9\linewidth]{images/present_previous_allplots.jpg}
\end{sidewaysfigure}
\clearpage

\textcolor{blue}{The Bayes factors for the syntactic interference effect show that the present study is the only one to provide evidence against the effect in the pre-critical region. In the critical region, both self-paced reading experiments provide evidence against the effect. The Bayes factors of three of the eye-tracking studies show weak or no evidence for the effect. The Bayes factor of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 2 provides extremely strong evidence for the syntactic interference effect in the critical region. In the spill-over region, the Bayes factors for syntactic interference rather consistently provide evidence against the effect. Only \citeauthor{mertzen}'s (\citeyear{mertzen}) German experiment provides moderate evidence for it.}

\textcolor{blue}{The Bayes factors for the semantic interference effect consistently provide evidence for the effect in the pre-critical region. In the critical region, most experiments provide evidence against semantic interference, the two self-paced reading experiments, being the exceptions, provide evidence for the effect. In the spill-over region, the Bayes factor results for semantic interference are mixed: the two self-paced reading experiments provide extremely strong evidence for the effect, the total fixation times for two of the eye-tracking experiments provide weak evidence for the effect and the total fixation times of two other eye-tracking experiments provide evidence against the effect. Importantly, the clustering in the total fixation time results does not align with the authors of the studies.}

\textcolor{blue}{The Bayes factors for the interaction in all regions are rather consistent between studies and are either inconclusive or provide evidence against the interaction. The only exception is the very large Bayes factor of \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 1. In sum, while the present and previous studies show comparable region-wise results for semantic interference and the interaction, they differ remarkably regarding syntactic interference in the pre-critical and critical region. In the pre-critical region, the present study is the obvious outlier, being the only study to provide evidence against the effect. In the critical region, four of the investigated experiments show weak evidence for or against syntactic interference with two outliers: \citeauthor{vandyke07}'s (\citeyear{vandyke07}) Experiment 2 providing extreme evidence for the effect and the present study providing strong evidence against it. In conclusion, the most remarkable difference between the present study and previous findings concerns the syntactic interference in the pre-critical region.}

%The difference in observed effects \textcolor{blue}{between our SPR experiment and the other reading time studies} could be explained by differences in statistical power or by methodological differences.

The present study has higher statistical power than the previous studies. \citet{vandyke07} had 35 - 40 participants and 36 - 48 items in each of her three experiments, respectively. \textcite{mertzen} tested 61 English speakers and 121 German speakers with 40 items each. In contrast, we tested 774 participants with at least 60 items each (a subset of 160 participants read all 120 items). Therefore, the present study is the study with the highest power to date using this 2 $\times$ 2 interference design. Consequently, the estimates from the present study are the most precise ones so far for this design.\textcolor{blue}{This is reflected by the tighter estimates in Figure \ref{fig:previous_vs_present}}. The smaller effect size in the present study is a consequence of higher power, and the magnitude of the effects observed is comparable to that of other large-scale reading studies \citep{nicenboim} and meta-analyses on interference \citep{jaeger_etal_2017}.

\Copy{syn_confounds}{\textcolor{blue}{The difference in methodology lead to another difference between our study and the previous studies. In their eye-tracking experiments, \textcite{vandyke07} and \textcite{mertzen} presented the stimuli on a single line. Whether \citeauthor{vandyke07}'s \citeyear{vandyke07} SPR experiment used line breaks or not is not reported. In our SPR experiments, it was necessary to introduce line breaks into the sentences. These were hard-coded to occur always after the subject of the critical verb and two words after the critical verb. The line break after the subject might have increased its prominence compared to the other nouns in the sentence, i.e., the distractor. It has been proposed by \textcite{engelmann_etal_2019} that prominence within the current context increases the activation of a memory representation which leads to faster retrieval and higher retrieval probability. Therefore, increased prominence of the subject could have decreased interference. Since we have found evidence for semantic, but not for syntactic interference, the increased prominence might have especially affected syntactic interference. For this point to hold, one would need to assume that the line break provided additional structural information by emphasizing the syntactic structure of the items in our SPR experiment. However, the EEG experiment used word-by-word presentation for all words, so there was no additional structural information which might have increased the prominence of the subject compared to the distractor. So, the line break might have decreased syntactic interference in our SPR experiments, but not in the EEG experiment. Nevertheless, neither our SPR nor our EEG experiments showed evidence for syntactic interference.}

\textcolor{blue}{All our experiments (SPR and EEG) displayed the subject with a comma. Commas are mandatory in German to separate the subordinate clauses from the main clause. Similar to the line break discussed above, the comma could have emphasized the syntactic structure of the sentences and therefore decreased syntactic interference from the distractor which occurred inside a relative clause and not in the clause of the critical verb. But if this comma provided information about the syntactic structure of the items then it should have so as well in \citeauthor{mertzen}'s \citeyear{mertzen} German study because they also presented the subject with the comma. So, while it is plausible that commas provide structural information, it cannot explain the difference in findings between our study and the previous studies.}}\label{syn_confounds}

So, why did we find a semantic interference effect in all regions, whereas \citet{mertzen} only found one in the pre-critical region? Apart from \textcolor{blue}{the additional distractor and} the methodological difference (eye-tracking vs.\ self-paced reading), a plausible  explanation for the divergence is lower power in \citeauthor{mertzen}'s (\citeyear{mertzen}) study. While they already had considerably more participants and items than many psycholinguistic studies, in the present study the number of participants was much higher than in theirs (774 vs.\ 121) and the number of items was also higher (60/120 vs.\ 40).
